\thetitle

### 2021-05-31 writing/discussion.tex
discussion: reword 'why transient lose blame'

An investigation of these empty blame cases reveals that Transient is not
unsound but that Vitousek et al.'s Transient blame assignment loses track
of the proper boundaries due to an incomplete initialization of the blame map.
Despite these losses, the number of failures is a small fraction of the
overall number of debugging scenarios. It is unlikely that any
improvements to Vitousek et al.'s strategy would change the overall
conclusion about Transient blame.

{\bf Note} The scenarios where Transient returns no blame suggest two ways to
improve its blame map.  First, entries in a blame map should point to {\em
several\/} parent entries.  For example, if \texttt{f} receives bad input in the
context of {\tt (filter f xs)}, the entry should point to both \texttt{filter}
and \texttt{xs}. Second, the construction of entries must be guided by type-like
specifications for primitives instead of special cases for built-in
functions. Aliasing such function changes the blame map.

>>>>

An investigation of these empty blame cases reveals problems with tracking blame
for higher-order functions, but conversely suggests three improvements for the
Transient algorithm. To illustrate, consider the call \texttt{(filter f xs)}.
The first suggestion is that the blame map should know that inputs to
\texttt{f} may have come from the \texttt{xs} list; concretely, the blame-map
entry for \texttt{f} should point to \texttt{xs} as a parent.
Second, there should be two parents for \texttt{f} instead of one, because
both \texttt{xs} and \texttt{filter} are responsible for sending correct input
to \texttt{f}. Third, the blame map should work equally well in programs
that rename \texttt{filter} or that replace the identifier with
an expression. This third point suggests a need for type-like specifications
that guide the construction of the blame map, instead of the identifier-based
matching in Reticulated~\cite{vss-popl-2017} and Shallow Racket.
